# Non-Communicable Disease Burden (NCD)

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data/burden-of-disease-ncd.csv"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Filter the dataset to include only the three countries chosen by your group.
data_ncd_mini <- data_ncd %>% 
  filter(Entity %in% c("Uganda", "Rwanda", "Kenya")) %>% 
  rename(DALYs_ncd = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(-Code)
```

## Table of Estimates for NCD Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_ncd_wide <- data_ncd_mini %>%
    pivot_wider(names_from  = Entity,
    values_from = DALYs_ncd)

kable(data_ncd_wide)
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in NCD Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
data_ncd_mini %>%
  ggplot(aes(x = Year,y = DALYs_ncd, color = Entity)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "DALYs (NCD)", color = "Country") +
  theme_minimal()
```

## Summary of NCD Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
- Uganda registeredhigh CNDs in the 1900 as compared to rwandan and Kenya.
- Ugandas NCD rate has dropped over the years as compared to Kenya and Rwanda.

## Attribution

This section was completed by [MWEBEMBEZI JOHN FISHER]

The link to the group's GitHub repository is [WRITE THE LINK HERE!]
