# Overall Disease Burden

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data/burden-of-disease-all-causes.csv"))
  
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Filter the dataset to include only the three countries chosen by your group.
data_overall_mini <-  data_overall %>% 
  filter(Entity %in% c("Uganda", "Rwanda", "Kenya")) %>% 
  rename(DALYs_overall = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(-Code)
```

## Table of Estimates for Overall Disease Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_overall_wide <- data_overall_mini %>%
    pivot_wider(names_from  = Entity,
    values_from = DALYs_overall)

kable(data_overall_wide)
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
data_overall_mini %>%
  ggplot(aes(x = Year,y = DALYs_overall, color = Entity)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "DALYs (Overall)", color = "Country") +
  theme_minimal()
```

## Summary of Overall Disease Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
- Overall the Uganda registered less cases of the disease since the 2000
- Rwanda has had a steady decline overtime

## Attribution

This section was completed by [MWEBEMBEZI JOHN FISHER]

The link to the group's GitHub repository is [WRITE THE LINK HERE!]
