---
title: "Global Disease Burden Analysis"
author: "REPLACE WITH THE AUTHOR NAMES" # AUTHOR NAMES HERE!
output: html_document
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

# Introduction

This document compares the Disability-Adjusted Life Year (DALY) diseases burden for three countries: COUNTRY 1, COUNTRY 2, and COUNTRY 3. We compare the burden due to communicable, maternal, neonatal, and nutritional diseases (CMNN), non-communicable diseases (NCDs), and overall disease burden for these countries.

The source data is from the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease (GBD) study.

```{r child= c('01_cmnn_burden_CHILD.Rmd', '02_ncd_burden_CHILD.Rmd', '03_overall_burden_CHILD.Rmd')}
```
# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data/burden-of-disease-cmnn.csv"))
```

```{r echo=FALSE, warning=FALSE}
# Filter the dataset to include only the three countries chosen by your group.
data_cmnn_mini <- data_cmnn %>% 
  filter(Entity %in% c("Uganda", "Rwanda", "Kenya")) %>% 
  rename(DALYs_cmmn = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(-Code)
```

## Table of Estimates for CMNN Burden Over Time

```{r echo=FALSE, warning=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_cmnn_wide <- data_cmnn_mini %>%
    pivot_wider(names_from  = Entity,
    values_from = DALYs_cmmn)

kable(data_cmnn_wide)
  
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in CMNN Burden Over Time

```{r echo=FALSE, warning=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.

data_cmnn_mini %>%
  ggplot(aes(x = Year,y = DALYs_cmmn, color = Entity)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "DALYs (CMMN)", color = "Country") +
  theme_minimal()
```

## Summary of CMNN Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
- DALY burden of the listed disease has gradually dropped over the years.
- Uganda has toppped all cases over the years.

# Non-Communicable Disease Burden (NCD)

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data/burden-of-disease-ncd.csv"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Filter the dataset to include only the three countries chosen by your group.
data_ncd_mini <- data_ncd %>% 
  filter(Entity %in% c("Uganda", "Rwanda", "Kenya")) %>% 
  rename(DALYs_ncd = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(-Code)
```

## Table of Estimates for NCD Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_ncd_wide <- data_ncd_mini %>%
    pivot_wider(names_from  = Entity,
    values_from = DALYs_ncd)

kable(data_ncd_wide)
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in NCD Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
data_ncd_mini %>%
  ggplot(aes(x = Year,y = DALYs_ncd, color = Entity)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "DALYs (NCD)", color = "Country") +
  theme_minimal()
```

## Summary of NCD Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
- Uganda registeredhigh CNDs in the 1900 as compared to rwandan and Kenya.
- Ugandas NCD rate has dropped over the years as compared to Kenya and Rwanda.

# Overall Disease Burden

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data/burden-of-disease-all-causes.csv"))
  
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Filter the dataset to include only the three countries chosen by your group.
data_overall_mini <-  data_overall %>% 
  filter(Entity %in% c("Uganda", "Rwanda", "Kenya")) %>% 
  rename(DALYs_overall = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(-Code)
```

## Table of Estimates for Overall Disease Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_overall_wide <- data_overall_mini %>%
    pivot_wider(names_from  = Entity,
    values_from = DALYs_overall)

kable(data_overall_wide)
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
data_overall_mini %>%
  ggplot(aes(x = Year,y = DALYs_overall, color = Entity)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "DALYs (Overall)", color = "Country") +
  theme_minimal()
```

## Summary of Overall Disease Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
- Overall the Uganda registered less cases of the disease since the 2000
- Rwanda has had a steady decline overtime

## Attribution

This section was completed by [MWEBEMBEZI JOHN FISHER]

The link to the group's GitHub repository is [WRITE THE LINK HERE!]
